### 2018-01-31 includes/experiment.tex
Adding more data to tables. Formatting tables for easier editing. Adding two more test problems.

\begin{figure}[h]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-extracts-small.png}
        \caption{Performance on concats-extracts-small}
        \label{fig:concats-extracts-small}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/different-prefix.png}
        \caption{Performance on different-prefix}
        \label{fig:different-prefix}
    \end{subfigure}
    \caption{Instances hard for \cvc{}}
    \label{fig:cvc-hard}

    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-balanced.png}
        \label{fig:concats-balanced}
        \caption{Performance on concats-balanced}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-small.png}
        \label{fig:concats-small}
        \caption{Performance on concats-small}
    \end{subfigure}
    \caption{Instances hard for \us{}}
    \label{fig:z3str3-hard}
\end{figure}

@@

\fuzzer{}'s ability to produce scaling instances (e.g., long string constants,
concats of increasing depth, etc.) helped uncover several
implementation issues and performance limitations in \us{}. Scaling inputs
can reveal issues that would be out of scope for unit tests or
industrial benchmarks because they elicit asymptotic behaviour. Three different
performance and implementation bugs were identified and fixed in \us{}
as a result of testing with the \fuzzer{} suites
\textit{Lengths} and \textit{Concats}.

>>>>

\begin{figure}[h]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-extracts-small.png}
        \caption{Performance on \textit{Concats-Extracts-Small}}
        \label{fig:concats-extracts-small}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/different-prefix.png}
        \caption{Performance on \textit{Different-Prefix}}
        \label{fig:different-prefix}
    \end{subfigure}
    \caption{Instances hard for \cvc{}}
    \label{fig:cvc-hard}

    \vspace{0.15in}

    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-balanced.png}
        \label{fig:concats-balanced}
        \caption{Performance on \textit{Concats-Balanced}}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-small.png}
        \label{fig:concats-small}
        \caption{Performance on \textit{Concats-Small}}
    \end{subfigure}
    \caption{Instances hard for \us{}}
    \label{fig:z3str3-hard}
\end{figure}

@@

\fuzzer{}'s ability to produce scaling instances helped uncover several
implementation issues and performance limitations in \us{}. Scaling inputs
can reveal issues that would be out of scope for unit tests or
industrial benchmarks because they elicit asymptotic behaviour. Three different
performance and implementation bugs were identified and fixed in \us{}
as a result of testing with the \fuzzer{} suites
\textit{Lengths} and \textit{Concats}.

### 2018-01-31 includes/stringfuzz.tex
Adding more data to tables. Formatting tables for easier editing. Adding two more test problems.

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string and regex
instances. It can either be installed from source, or from the Python
PIP package repository.\footnote{The link to the
source code will be added after double-blind review.}
\fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. Its
components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
        This tool generates \smt{} instances. It supports several generators and
        options that specify its output. Details can be found in
        table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
        This tool transforms \smt{}
        instances. It supports several transformers and options that specify
        its output and input, which are explained in
        table~\ref{tbl:transformers}. Note that transformers
        \textit{Translate} and \textit{Reverse} also preserve
        satisfiability under certain conditions~\cite{ifaz}.
    \item[\texttt{stringstats}] \hfill \\
        This tool takes an \smt{}
        instance as input and outputs its properties: the number of
        variables/literals, the max/median syntactic depth of expressions, the
        max/median literal length, etc.
\end{description}

\begin{table}[t]
    \centering
    \caption{\generator{} generators.}
    \label{tbl:generators}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{Generated instances have ...} \\ \hline
        \textit{Concats}     & Long concats and optional random extracts. \\ \hline
        \textit{Lengths}     & Many variables (and their concats) with length constraints. \\ \hline
        \textit{Overlaps}    & An expression of the form A.X = X.B. \\ \hline
        \textit{Equality}    & An equality among concats, each with variables or constants. \\ \hline
        \textit{Regex}       & Regexes of varying complexity. \\ \hline
        \textit{Random-Text} & Random, likely syntactically \textit{in}valid text. \\ \hline
        \textit{Random-AST}  & Random, but semantically \textit{valid} text. \\ \hline
    \end{tabular}
    \caption{\transformer{} transformers.}
    \label{tbl:transformers}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{The transformer ...} \\ \hline
        \textit{Fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
        \textit{Graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
        \textit{Multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
        \textit{Nop}         & Doesn't transform (but can translate \smtfull{}). \\ \hline
        \textit{Reverse}     & Reverses all string literals and concat arguments. \\ \hline
        \textit{Rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
        \textit{Translate}   & Maps every character X to a different character Y. \\ \hline
        \textit{Unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
    \end{tabular}
\end{table}

@@

\subsubsection{Example Use Cases}

\fuzzer{} can be used to
generate benchmark suites in a batch mode, or can be used in an online
randomized fashion for exploratory debugging. For example, \fuzzer{}
can be invoked as follows to feed random instances to \cvc{} until it
returns an error:
% \vspace{-0.15in}
% \begin{figure}[H]
%     \centering
%     \begin{minipage}[b]{0.91\linewidth}
%         \begin{minted}{bash}
{\scriptsize\begin{verbatim}
while stringfuzzg -r random-ast -m | tee instance.smt25 | cvc4 --lang smt2.5 --tlimit=5000; do
    sleep 0
done\end{verbatim}}
%         \end{minted}
%     \end{minipage}
% \end{figure}
% \vspace{-0.15in}

Some more \unix{} scripting can yield other interesting results. For
example, the following script finds a regex instance on which \us{}
times out after 5 seconds:
% \vspace{-0.15in}
% \begin{figure}[H]
%     \centering
%     \begin{minipage}[b]{0.91\linewidth}
%         \begin{minted}{bash}
{\scriptsize\begin{verbatim}while true; do
    stringfuzzg -r regex -c -a increasing -r 2 -t 5 -d 3 > instance.smt25
    result=`z3str3 -T:5 instance.smt25`
    echo $result
    if [ "$result" == "timeout" ]; then
        cat instance.smt25
        break
    fi
done\end{verbatim}}
%         \end{minted}
%     \end{minipage}
% \end{figure}
% \vspace{-0.15in}

>>>>

\subsubsection{Implementation and Architecture}

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string and regex
instances. It can either be installed from source, or from the Python
PIP package repository.\footnote{The link to the
source code will be added after double-blind review.}
\fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. To demonstrate
its modularity, we note that the implementations of transformers in \fuzzer{}
contain an average of \linesPerX{} lines of code, whereas the whole fuzzer
contains \linesInFuzzer{} lines.

Its components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
        This tool generates \smt{} instances. It supports several generators and
        options that specify its output. Details can be found in
        table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
        This tool transforms \smt{}
        instances. It supports several transformers and options that specify
        its output and input, which are explained in
        table~\ref{tbl:transformers}. Note that transformers
        \textit{Translate} and \textit{Reverse} also preserve
        satisfiability under certain conditions~\cite{ifaz}.
    \item[\texttt{stringstats}] \hfill \\
        This tool takes an \smt{}
        instance as input and outputs its properties: the number of
        variables/literals, the max/median syntactic depth of expressions, the
        max/median literal length, etc.
\end{description}

\begin{table}[t]
    \centering
    \caption{\generator{} generators.}
    \label{tbl:generators}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}
            & \textbf{Generated instances have ...}
            \\ \hline
        \textit{Concats}
            & Long concats and optional random extracts.
            \\ \hline
        \textit{Lengths}
            & Many variables (and their concats) with length constraints.
            \\ \hline
        \textit{Overlaps}
            & An expression of the form A.X = X.B.
            \\ \hline
        \textit{Equality}
            & An equality among concats, each with variables or constants.
            \\ \hline
        \textit{Regex}
            & Regexes of varying complexity.
            \\ \hline
        \textit{Random-Text}
            & Random, likely syntactically \textit{in}valid text.
            \\ \hline
        \textit{Random-AST}
            & Random, but semantically \textit{valid} text.
            \\ \hline
    \end{tabular}

    \caption{\transformer{} transformers.}
    \label{tbl:transformers}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}
            & \textbf{The transformer ...}
            \\ \hline
        \textit{Fuzz}
            & Randomly replaces literals and operators with similar ones.
            \\ \hline
        \textit{Graft}
            & For each sort, swaps a random non-leaf node with a leaf node.
            \\ \hline
        \textit{Multiply}
            & Multiplies each number by N, and repeats each character N times.
            \\ \hline
        \textit{Nop}
            & Doesn't transform (but can translate \smtfull{}).
            \\ \hline
        \textit{Reverse}
            & Reverses all string literals and concat arguments.
            \\ \hline
        \textit{Rotate}
            & Performs a random syntax tree rotation on two concats.
            \\ \hline
        \textit{Translate}
            & Maps every character X to a different character Y.
            \\ \hline
        \textit{Unprintable}
            & Randomly replaces all characters in literals with unprintable ones.
            \\ \hline
    \end{tabular}
\end{table}

@@

\subsubsection{Example Use Case}

\fuzzer{} can be used to generate benchmark suites in a batch mode, but also
in an online randomized fashion for exploratory debugging. For example,
\fuzzer{} can be invoked as follows to feed random instances to \cvc{} until it
returns an error:
{\scriptsize\begin{verbatim}
while stringfuzzg -r random-ast -m | tee instance.smt25 | cvc4 --lang smt2.5 --tlimit=5000; do
    sleep 0
done\end{verbatim}}

Simply by running the above script, we managed to induce segmentation faults
in \cvc{} and \us{}.
